Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-19 paper/main.tex
new figures added

Ekhine: How many runs do we do? 10 20?

To better understand the behaviour of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over X runs on the same problem instances.  As an example, Fig.~\ref{fig:kra32}
shows the mean fitness of the solutions evaluated by uMM and CEGO on one QAP
instance. The shaded area shows one standard deviation around the
mean. Ideally, each solution evaluated helps to refine the model and leads to a
better solution being evaluated in the next step.  However, the plot shows that
the fitness of the solutions does not improve monotonically and new solutions
are often worse than their predecessors. Nevertheless, in this example, there
is a clear overall improvement as more evaluations are performed. The minimum
values achieved in the plot up to a particular number of function evaluations
gives an estimation of the mean fitness of the best-so-far solution up to that
point in the run. The algorithms do not return the final solution shown but
rather the best one (minimum fitness) of each run.

@@

The fast initial convergence of CEGO for synthetic LOP instances does not seem
to translate to the instances available in LOPLIB~\citep{}. We show in
Fig.~\ref{fig:loplib} three examples with different permutation size, however,
the results are consistent for other LOPLIB instances (complete results
available as supplementary material \supplement). Due to extremely long
runtimes of CEGO and constraints in our computing system, we set a maximum
wall-clock time of 5 days for each run of CEGO. On these instances, the
behaviour of CEGO and uMM is much more similar (at least up to the termination
point of CEGO). It appears that our synthetic LOP instances have a fitness
landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
instances do not present such landscape. Comparing CEGO and uMM results, the
differences get smaller, in favour of uMM, with larger instance size. Our
conjecture is that building an accurate Bayesian model and searching for good
solutions on it becomes harder for larger fitness landscapes.

@@

\begin{table}
  \caption{Results.\label{tab:results}}
%\begin{adjustbox}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lllrrrrrrr}
\toprule
{} & Problem &                instance &  Fitness\_CEGO\_mean &  Fitness\_CEGO\_std &  Fitness\_UMM\_mean &  Fitness\_UMM\_std &  Improvement &  ttest statistic &  ttest pvalue \\
\midrule
0  &     LOP &  lop/RandA1/N-t1d100.01 &            78696.6 &        811.368145 &           76119.6 &       915.392836 &     0.032746 &         6.662091 &  2.989433e-06 \\
1  &     LOP &  lop/RandA1/N-t1d100.02 &            79686.5 &        367.876262 &           76827.7 &       890.969391 &     0.035876 &         9.378614 &  2.374606e-08 \\
2  &     LOP &  lop/RandA1/N-t1d150.01 &           187605.5 &       1906.941953 &          179508.3 &      1647.878437 &     0.043161 &        10.159726 &  6.994357e-09 \\
3  &     LOP &  lop/RandA1/N-t1d150.02 &           184320.9 &        973.457920 &          177592.8 &      1196.159298 &     0.036502 &        13.795848 &  5.183671e-11 \\
4  &     LOP &  lop/RandA1/N-t1d200.01 &           335391.1 &       2288.960096 &          323513.6 &      1371.235712 &     0.035414 &        14.076557 &  3.714220e-11 \\
5  &     LOP &  lop/RandA1/N-t1d200.02 &           333149.5 &       2497.207474 &          319675.8 &      2242.631629 &     0.040443 &        12.694417 &  2.029610e-10 \\
6  &     LOP &  lop/RandA2/N-t2d150.01 &            24682.7 &        976.639596 &           15297.7 &       563.527787 &     0.380226 &        26.320561 &  8.038269e-16 \\
7  &     LOP &  lop/RandA2/N-t2d150.02 &            24250.8 &       1354.897110 &           14724.9 &       667.454276 &     0.392808 &        19.944373 &  1.010663e-13 \\
8  &    PFSP &          pfsp/rec05.txt &             1317.9 &         25.075220 &            1330.7 &        17.901272 &    -0.009712 &        -1.313790 &  2.054160e-01 \\
9  &    PFSP &          pfsp/rec13.txt &             2128.2 &         24.965754 &            2158.7 &        29.533597 &    -0.014331 &        -2.494040 &  2.258601e-02 \\
10 &    PFSP &          pfsp/rec19.txt &             2398.6 &         19.839355 &            2405.1 &        12.582616 &    -0.002710 &        -0.874932 &  3.931359e-01 \\
11 &     QAP &           qap/kra32.dat &           115907.0 &       4663.427566 &          116320.0 &      4673.468852 &    -0.003563 &        -0.197816 &  8.454055e-01 \\
12 &     QAP &           qap/nug12.dat &              624.2 &         26.288147 &             660.4 &        17.276831 &    -0.057994 &        -3.639054 &  1.876810e-03 \\
13 &     QAP &           qap/nug30.dat &             7486.8 &         40.411220 &            7474.0 &        83.910005 &     0.001710 &         0.434612 &  6.690098e-01 \\
14 &     QAP &           qap/tho30.dat &           192416.0 &       2584.174401 &          191482.0 &      4419.722264 &     0.004854 &         0.576896 &  5.711548e-01 \\
\bottomrule
\end{tabular}}
% \end{adjustbox}
 \end{table}

@@

\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
\bibliographystyle{splncs04nat}
\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref}

>>>>

Ekhine: How many runs do we do? 10 20? 10, cambio el texto

To better understand the behaviour of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs on the same problem instances.  As an example, Fig.~\ref{fig:kra32} \textcolor{red}{por qué no mepezamos con la fig1?}
shows the mean fitness of the solutions evaluated by uMM and CEGO on one QAP
instance. The shaded area shows one standard deviation around the
mean. Ideally, each solution evaluated helps to refine the model and leads to a
better solution being evaluated in the next step.  However, the plot shows that \textcolor{red}{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}
the fitness of the solutions does not improve monotonically and new solutions
are often worse than their predecessors. Nevertheless, in this example, there
is a clear overall improvement as more evaluations are performed. The minimum
values achieved in the plot up to a particular number of function evaluations
gives an estimation of the mean fitness of the best-so-far solution up to that
point in the run. The algorithms do not return the final solution shown but
rather the best one (minimum fitness) of each run.

@@

The fast initial convergence of CEGO for synthetic LOP instances does not seem
to translate to the instances available in LOPLIB~\citep{}. We show in
Fig.~\ref{fig:loplib} three examples with different permutation size, however,
the results are consistent for other LOPLIB instances (complete results
available as supplementary material \supplement). Due to extremely long
runtimes of CEGO and constraints in our computing system, we set a maximum
wall-clock time of 5 days for each run of CEGO. On these instances, the
behaviour of CEGO and uMM is much more similar (at least up to the termination
point of CEGO). It appears that our synthetic LOP instances have a fitness
landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
instances do not present such landscape. Comparing CEGO and uMM results, the
differences get smaller, in favour of uMM, with larger instance size. Our
conjecture is that building an accurate Bayesian model and searching for good
solutions on it becomes harder for larger fitness landscapes. 
\textcolor{red}{Parece que el paper va del CEGO en vez del UMM :p. Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}

@@

%
%    \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
%\begin{table}
%  \caption{Results.\label{tab:results}}
%  \resizebox{\textwidth}{!}{%
%\begin{tabular}{*{6}{r}c*{3}{r}}
%  \toprule
%            &                  & \mcolc{4}{Mean fitness (sd)} &                 & \mcolc{2}{Mean runtime}             & \bf     Mean FE\\\cmidrule{3-6}\cmidrule(lr){8-9}
%\bf Problem & \bf     Instance & \mcolc{2}{UMM}               & \mcolc{2}{CEGO} & \bf 95\% CI for the mean difference & \bf UMM  & \bf CEGO                                  & \bf CEGO           \\\midrule
%    LOP     & N-t1d100.01      & 76119.6                      & (915.4)         & 78696.6                             & (811.4)  & (1763.4904574743996, 3390.5095425256004)  & 11436.2 & 260954.5 & 111.8 \\
%    LOP     & N-t1d100.02      & 76827.7                      & (891.0)         & 79686.5                             & (367.9)  & (2194.5410893080507, 3523.058910691955)   & 11353.3 & 262421.6 & 110.7 \\
%    LOP     & N-t1d150.01      & 179508.3                     & (1647.9)        & 187605.5                            & (1906.9) & (6420.259663264915, 9774.140336735109)    & 23395.6 & 261712.4 & 73.9  \\
%    LOP     & N-t1d150.02      & 177592.8                     & (1196.2)        & 184320.9                            & (973.5)  & (5700.461476832616, 7755.738523167395)    & 23361.2 & 263389.9 & 74.1  \\
%    LOP     & N-t1d200.01      & 323513.6                     & (1371.2)        & 335391.1                            & (2289.0) & (10076.07401989604, 13678.92598010396)    & 38476.8 & 263467.5 & 55.6  \\
%    LOP     & N-t1d200.02      & 319675.8                     & (2242.6)        & 333149.5                            & (2497.2) & (11241.971680405779, 15705.428319594244)  & 38268.3 & 265217.5 & 56.5  \\
%    LOP     & N-t2d150.01      & 15297.7                      & (563.5)         & 24682.7                             & (976.6)  & (8622.206776092156, 10147.793223907844)   & 25318.6 & 262324.3 & 74.6  \\
%    LOP     & N-t2d150.02      & 14724.9                      & (667.5)         & 24250.8                             & (1354.9) & (8495.057408450935, 10556.742591549064)   & 25425.6 & 262079.2 & 74.8  \\
%    LOP     & N-t2d200.01      & 33465.2                      & (2001.0)        & 57098.6                             & (4452.8) & (20284.508914101127, 26982.291085898876)  & 40340.9 & 263908.9 & 56.3  \\
%    LOP     & N-t2d200.02      & 35901.8                      & (2077.5)        & 58123.0                             & (2334.9) & (9515.181030361815, 34927.21896963818)    & 40153.1 & 261723.6 & 55.0  \\\midrule
%   PFSP     & rec05            & 1330.7                       & (17.9)          & 1317.9                              & (25.1)   & (-33.424760001214985, 7.8247600012150755) & 506.1   & 139893.0 & 400.0 \\
%   PFSP     & rec13            & 2158.7                       & (29.5)          & 2128.2                              & (25.0)   & (-56.24362499249925, -4.756375007500754)  & 506.8   & 140167.2 & 400.0 \\
%   PFSP     & rec19            & 2405.1                       & (12.6)          & 2398.6                              & (19.8)   & (-22.313869704790818, 9.313869704790818)  & 1078.3  & 295442.7 & 400.0 \\\midrule
%    QAP     & kra32            & 116320.0                     & (4673.5)        & 115907.0                            & (4663.4) & (-4799.295389697205, 3973.295389697205)   & 1194.8  & 344766.2 & 400.0 \\
%    QAP     & nug12            & 660.4                        & (17.3)          & 624.2                               & (26.3)   & (-57.337505259978144, -15.06249474002172) & 180.0   & 53149.3  & 400.0 \\
%    QAP     & nug30            & 7474.0                       & (83.9)          & 7486.8                              & (40.4)   & (-50.84533806724974, 76.4453380672501)    & 1073.5  & 303352.6 & 400.0 \\
%    QAP     & tho30            & 191482.0                     & (4419.7)        & 192416.0                            & (2584.2) & (-2527.0213832396516, 4395.021383239651)  & 1066.5  & 306451.3 & 400.0 \\
%\bottomrule
%\end{tabular}}
%\end{table}

  \resizebox{\textwidth}{!}{
\begin{tabular}{lllllrrr}
\toprule
Problem &       Instance &           UMM fitness &           CEGO fitness &     Confidence interval &  UMM Runtime &  CEGO Runtime &  CEGO F. evaluations \\
\midrule
    LOP &    N-t1d100.01 &       76119.6 (915.4) &        78696.6 (811.4) &        (1763.5, 3390.5) &          3.2 &          72.5 &                111.8 \\
    LOP &    N-t1d100.02 &       76827.7 (891.0) &        79686.5 (367.9) &        (2194.5, 3523.1) &          3.2 &          72.9 &                110.7 \\
    LOP &    N-t1d150.01 &     179508.3 (1647.9) &      187605.5 (1906.9) &        (6420.3, 9774.1) &          6.5 &          72.7 &                 73.9 \\
    LOP &    N-t1d150.02 &     177592.8 (1196.2) &       184320.9 (973.5) &        (5700.5, 7755.7) &          6.5 &          73.2 &                 74.1 \\
    LOP &    N-t1d200.01 &     323513.6 (1371.2) &      335391.1 (2289.0) &      (10076.1, 13678.9) &         10.7 &          73.2 &                 55.6 \\
    LOP &    N-t1d200.02 &     319675.8 (2242.6) &      333149.5 (2497.2) &      (11242.0, 15705.4) &         10.6 &          73.7 &                 56.5 \\
    LOP &    N-t2d150.01 &       15297.7 (563.5) &        24682.7 (976.6) &       (8622.2, 10147.8) &          7.0 &          72.9 &                 74.6 \\
    LOP &    N-t2d150.02 &       14724.9 (667.5) &       24250.8 (1354.9) &       (8495.1, 10556.7) &          7.1 &          72.8 &                 74.8 \\
    LOP &    N-t2d200.01 &      33465.2 (2001.0) &       57098.6 (4452.8) &      (20284.5, 26982.3) &         11.2 &          73.3 &                 56.3 \\
    LOP &    N-t2d200.02 &      35901.8 (2077.5) &       54895.6 (4589.7) &      (15539.1, 22448.5) &         11.2 &          73.5 &                 55.6 \\
    LOP &       N-p50-01 &       21027.6 (711.8) &        21431.2 (299.5) &         (-128.1, 935.3) &          0.9 &          72.3 &                219.9 \\
    LOP &       N-p50-02 &       21613.6 (596.4) &        22021.1 (275.6) &          (-42.5, 857.5) &          0.9 &          72.3 &                217.3 \\
    LOP &       N-atp111 &          594.0 (15.4) &           693.1 (10.3) &           (86.7, 111.5) &          4.1 &          72.5 &                 98.7 \\
    LOP &       N-atp134 &          756.3 (21.3) &           880.5 (18.6) &          (105.4, 143.0) &          5.7 &          72.8 &                 83.0 \\
    LOP &  N-be75eec\_150 &   1415208.5 (58593.5) &    1666193.1 (72310.8) &    (188960.6, 313008.6) &          6.9 &          72.9 &                 74.0 \\
    LOP &  N-be75eec\_250 &  4431271.0 (109756.7) &   5214413.0 (166701.4) &    (649040.8, 917243.2) &         15.5 &          73.7 &                 44.9 \\
    LOP &   N-be75np\_150 &   3204878.3 (80704.1) &    3742444.0 (88500.2) &    (457944.2, 617187.2) &          6.9 &          72.9 &                 74.2 \\
    LOP &   N-be75np\_250 &  9544754.9 (260523.9) &  10887427.8 (159642.5) &  (1136635.8, 1548710.0) &         15.5 &          74.0 &                 45.3 \\
   PFSP &      rec05.txt &         1330.7 (17.9) &          1317.9 (25.1) &            (-33.4, 7.8) &          0.1 &          38.9 &                400.0 \\
   PFSP &      rec13.txt &         2158.7 (29.5) &          2128.2 (25.0) &           (-56.2, -4.8) &          0.1 &          38.9 &                400.0 \\
   PFSP &      rec19.txt &         2405.1 (12.6) &          2398.6 (19.8) &            (-22.3, 9.3) &          0.3 &          82.1 &                400.0 \\
    QAP &      kra32.dat &     116320.0 (4673.5) &      115907.0 (4663.4) &       (-4799.3, 3973.3) &          0.3 &          95.8 &                400.0 \\
    QAP &      nug12.dat &          660.4 (17.3) &           624.2 (26.3) &          (-57.3, -15.1) &          0.0 &          14.8 &                400.0 \\
    QAP &      nug30.dat &         7474.0 (83.9) &          7486.8 (40.4) &           (-50.8, 76.4) &          0.3 &          84.3 &                400.0 \\
    QAP &      tho30.dat &     191482.0 (4419.7) &      192416.0 (2584.2) &       (-2527.0, 4395.0) &          0.3 &          85.1 &                400.0 \\
\bottomrule
\end{tabular}}

@@

\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
\bibliographystyle{splncs04nat}
\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref,./mendeley}
